Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions

### 2010-11-29 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@41 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The most popular nonparametric goodness-of-fit test is the
Kolmogorov-Smirnov test.
Given the cumulative distribution
function $F_0(x)$ of the hypothesized distribution
and the empirical distribution function $F_{data}(x)$ of the
observed data,  the test statistic is given by
\begin{align}
D = \sup_x \left| F_0(x)- F_{data}(x) \right|.    \label{eqD}
\end{align}
When $F_0$ is continuous, 
the distribution of $D$ does not depend on the hypothesized
distribution, making this a computationally
attractive method. \cite{slakter1965} offers a standard presentation
of the test and its
performance relative to other algorithms. 
The test statistic is easily adapted for one-sided tests.
For these, the absolute value in (\ref{eqD}) is discarded and the tests are based
on either the supremum of the remaining difference (the `greater' testing
alternative) or by replacing the supremum with a negative infimum
(the `lesser' hypothesis alternative).  Tabulated p-values have been
available for these tests since 1933.  **** CITATION? ***

@@

While the Kolmogorov-Smirnov test may be the most popular of
the nonparametric goodness-of-fit tests, Cram\'{e}r-von Mises
tests have been shown to be more powerful against a large class
of alternatives hypotheses. 
The original test was developed by
Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
The original test statistic, $W^2$, Anderson's $A^2$, and Watson's
$U^2$ are:
\begin{align}
W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \label{W2} \\
A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \label{A2} \\
U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x) \label{U2}
\end{align}
As with the original Kolmogorov-Smirnov test statistic, these all have 
test statistic null distributions which are independent of the
hypothesized continuous models.

The $W^2$ statistic was the original test statistic.
The $A^2$ statistic was developed by
Anderson in the process of generalizing the test for the two-sample case.
Watson's $U^2$ statistic was developed for distributions
which are cyclic (with an ordering to the support but
no natural starting point); it is invariant to cyclic reordering of
the support.   For example, a distribution on the months of the year
could be considered cyclic.

It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations from the hypothesized distribution. 
They all involve integration over the whole range of data, rather than
use of a supremum, so they are best-suited for situations where the
true alternative distribution deviates a little over the whole support
rather than having large deviations over a small section of the support. 
\cite{stephens1974} offers a complete analysis of the relative powers
of these tests.

@@

data:  rep(1, 3) 
D = 0.6667, p-value = 0.04938
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
If, instead, either \code{exact=FALSE} is used with the new
\code{ks.test()} function, or if the original \code{stats::ks.test()}
is used, the reported p-value is 0.1389.

>>>>

The most popular nonparametric goodness-of-fit test is the
Kolmogorov-Smirnov test.
Given the cumulative distribution
function $F_0(x)$ of the hypothesized distribution
and the empirical distribution function $F_{data}(x)$ of the
observed data,  the test statistic is given by
\begin{align}
D = \sup_x \left| F_0(x)- F_{data}(x) \right|.    \label{eqD}
\end{align}
When $F_0$ is continuous, 
the distribution of $D$ does not depend on the hypothesized
distribution, making this a computationally
attractive method. \cite{slakter1965} offers a standard presentation
of the test and its
performance relative to other algorithms. 
The test statistic is easily adapted for one-sided tests.
For these, the absolute value in (\ref{eqD}) is discarded and the tests are based
on either the supremum of the remaining difference (the `greater' testing
alternative) or by replacing the supremum with a negative infimum
(the `lesser' hypothesis alternative).  Tabulated p-values have been
available for these tests since 1933 \citep{kol33}.

@@

While the Kolmogorov-Smirnov test may be the most popular of
the nonparametric goodness-of-fit tests, Cram\'{e}r-von Mises
tests have been shown to be more powerful against a large class
of alternatives hypotheses. 
The original test was developed by
Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
The original test statistic, $W^2$, Anderson's $A^2$, and Watson's
$U^2$ are:
\begin{align}
W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \label{W2} \\
A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \label{A2} \\
U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x) \label{U2}
\end{align}
As with the original Kolmogorov-Smirnov test statistic, these all have 
test statistic null distributions which are independent of the
hypothesized continuous models.
The $W^2$ statistic was the original test statistic.
The $A^2$ statistic was developed by
Anderson in the process of generalizing the test for the two-sample case.
Watson's $U^2$ statistic was developed for distributions
which are cyclic (with an ordering to the support but
no natural starting point); it is invariant to cyclic reordering of
the support.   For example, a distribution on the months of the year
could be considered cyclic.

It has been shown that these tests can be more powerful than 
Kolmogorov-Smirnov tests to certain deviations from the hypothesized distribution. 
They all involve integration over the whole range of data, rather than
use of a supremum, so they are best-suited for situations where the
true alternative distribution deviates a little over the whole support
rather than having large deviations over a small section of the support. 
\cite{stephens1974} offers a comprehensive analysis of the relative powers
of these tests.

@@

data:  rep(1, 3) 
D = 0.6667, p-value = 0.04938
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
If, instead, either \code{exact=FALSE} is used with the new
\code{ks.test()} function, or if the original \code{stats::ks.test()}
is used, the reported p-value is 0.1389 even though the test statistic
is the same.
